---
title: "Rainfall and Temperature"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: 
      version: 4
      bootswatch: minty
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(tidytuesdayR)
library(here)
library(ggplot2)
library(shiny)
library(DT)
```

# About

## Column { .tabset}

Rainfall and temperature are two important determining factors in agriculture production. However, factors like geographical region and interdependence between these two variables makes it hard to accurately model their relationship. 

Moreover, it is important to remind the mutual relationship between environmental factors (rainfall and temperature) and human activities (agricultural production). While environmental factors play a great role in determining crop production, modern agriculture also brings great impact to our ecosystem by accelerating climate change.

The goal of this dashboard is to help users explore the relationship between rainfall and temperature. 

<iframe
    width="640"
    height="480"
    src="https://www.youtube.com/embed/UmX4kyB2wfg"
    frameborder="0"
    allow="autoplay; encrypted-media"
    allowfullscreen
>
</iframe>

# Data

## Column { .tabset}

### Introduction to datasets 
The two datasets comes from [TidyTuesday](https://www.tidytuesday.com). Specifically, I will use data that comes from January 2020.

Here are the links to the original datasets: [rainfall](https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv) and [temperature](https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/temperature.csv)

If you want to find out what the column names represent, here is the link to the [data dictionary](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-07/readme.md#data-dictionary)

```{r, echo=FALSE}
if(!dir.exists(here("data"))){
  dir.create(here("data"))
}
if (!file.exists(here("data", "tuesdata_rainfall.RDS"))) {
    tuesdata <- tidytuesdayR::tt_load("2020-01-07")
    rainfall <- tuesdata$rainfall
    temperature <- tuesdata$temperature
    saveRDS(tuesdata$rainfall, file = here("data", "tuesdata_rainfall.RDS"))
    saveRDS(tuesdata$temperature, file = here("data", "tuesdata_temperature.RDS"))
}
rainfall <- readRDS(here("data", "tuesdata_rainfall.RDS"))
temperature <- readRDS(here("data", "tuesdata_temperature.RDS"))
```

## Column { .tabset .tab-fade}

### Rainfall
```{r}
DT::renderDataTable({
  DT::datatable(rainfall,
                caption = htmltools::tags$caption(
                  style = 'caption-side: top; text-align: Left;'),
                options = list(autoWidth = TRUE,
                               pageLength = 10,
                               scroller = TRUE,
                               scrollY = '450px'))
})
```

### Temperature
```{r}
DT::renderDataTable({
  DT::datatable(temperature,
                caption = htmltools::tags$caption(
                  style = 'caption-side: top; text-align: Left;'),
                options = list(autoWidth = TRUE,
                               pageLength = 10,
                               scroller = TRUE,
                               scrollY = '450px'))
})
```

# Interactive Analysis 

## Column {.sidebar}

```{r}
# user input (city)
selectInput(inputId = "input_city", 
              label = "Select a city:",
              choices = c("PERTH", "SYDNEY", "BRISBANE", "CANBERRA", "MELBOURNE"), 
            selected = "PERTH")
# user input (year)
numericInput(inputId = "input_year", 
             label = "Select a year:", 
             value = 2014, min = 1967, max = 2019)
sliderInput(inputId = "num_bin",
            label = "Number of bins:",
            min = 10,
            max = 30,
            value = 20)
```

## Distributions  { .tabset}

### Rainfall
```{r}
# data wrangling
rainfall_wrangled <- rainfall %>%
  na.omit(rainfall) %>%
  unite("date", c("year", "month", "day"), sep = "-", remove = FALSE) %>%
  mutate(date = ymd(date)) %>%
  select(-month, -day) %>%
  mutate(city_name = str_to_upper(city_name))
combined_data <- inner_join(x = rainfall_wrangled, y = temperature,  by=c('date', 'city_name'), relationship = "many-to-many")
# interactive plot
renderPlot({
  combined_data %>%
    filter(city_name == input$input_city, year == input$input_year) %>%
    ggplot(aes(log(rainfall))) +
    geom_histogram(colour = 4, fill = "white", bins = input$num_bin) +
    labs(title = paste("Distribution of rainfall (log scale) in", str_to_lower(input$input_city), "in", input$input_year),
         x = "Log-transformed rainfall",
         y = "count of rainfall amount") +
         theme(plot.title = element_text(hjust = 0.5, face = "bold"))
})
```

### Temperature
```{r}
renderPlot({
  combined_data %>%
    filter(city_name == input$input_city, year == input$input_year) %>%
    ggplot(aes(x = date, y = temperature, colour = temp_type)) +
    geom_line(size = 2) +
    labs(title = paste("Distribution of temperature in", str_to_lower(input$input_city), "in", input$input_year),
       x = "Date", y = "Temperature") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.text.x = element_text(hjust = 1, vjust = 1))
})
```

# Static Analysis

## Temperature change since 2014 { .tabset}

### Max
```{r}
combined_data %>%
  filter(year >= 2014, temp_type == "max") %>%
  ggplot(aes(x = date, y = temperature)) +
  geom_line(color = "pink") +
  facet_wrap(vars(city_name), ncol = 2) +
  labs(title = "Change of max temperature over time for each city since 2014",
       x = "Date", y = "Max temperature",
       caption = "Jiaxin Huang") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

### Min
```{r}
combined_data %>%
  filter(year >= 2014, temp_type == "min") %>%
  ggplot(aes(x = date, y = temperature)) +
  geom_line(color = "orange") +
  facet_wrap(vars(city_name), ncol = 2) +
  labs(title = "Change of min temperature over time for each city since 2014",
       x = "Date", y = "Min temperature",
       caption = "Jiaxin Huang") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

## Rainfall change since 2014 {data-height=700}

```{r}
calculate_CI <- function(x, conf) {
  alpha <- 1 - conf
  degrees_freedom <- length(x) - 1
  t_score <- qt(p = alpha / 2, df = degrees_freedom, lower.tail = FALSE)
  sample_sd_err <- sd(x)/sqrt(length(x))
  lower_bound <- mean(x) - t_score * sample_sd_err
  upper_bound <- mean(x) + t_score * sample_sd_err
  output_value <- c(lower_bound, upper_bound)
  output_name <- c("lower_bound", "upper_bound")
  output <- setNames(output_value, output_name)
  return(output)
}
rain_df <- combined_data %>%
  filter(year >= 2014) %>%
  group_by(year, city_name) %>%
  summarize(sample_mean = mean(rainfall),
            sample_sd = sd(rainfall),
            lower_bound = calculate_CI(rainfall, 0.95)["lower_bound"],
            upper_bound = calculate_CI(rainfall, 0.95)["upper_bound"])
rain_df %>%
  ggplot(aes(x = year, y = sample_mean, colour = "pink")) +
  geom_point(size = 2) +
  geom_line(colour = "black") +
  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound)) +
  facet_wrap(vars(city_name)) +
  labs(title = "Mean and 95% CI for rainfall over time for each city", 
       subtitle = "Canberra and Sydney have the lowest amount of rainfall, while the rest have relatively more rainfall.\nThe amount of rainfall tends to decrease as time goes by.",
       x = "Year", y = "Amount of rainfall",
       caption = "Jiaxin Huang") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle=element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

# Analysis

## Column

**Question**

How to explore the relationship between rainfall and temperature of a city in a given year and?
Here, I'll use the data for Perth in 2014 as an example.

**Target Audience**

1.  Any environmental specialists who are interested in exploring the relationship between temperature and rainfall, and use the relationship to predict possible health impacts on the population.
2.  Farmers who want to make plans for their crop production and develop risk reducing strategies.

**Methods**

First, let's view the raw distribution of rainfall data and temperature data for Perth in 2014:

```{r}
combined_data %>%
  filter(city_name == "PERTH", year == 2014) %>%
  ggplot(aes(x = date, y = rainfall)) +
  geom_point() +
  geom_line(aes(x = date, y = temperature, colour = temp_type)) +
  labs(title = "Distribution of temperature and rainfall for Perth in 2014", 
       caption = "Jiaxin Huang") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(hjust = 1, vjust = 1))
```

Then, let's view the relationship between rainfall and max temperature:

```{r}
combined_data %>%
  filter(city_name == "PERTH", year == 2014, temp_type == "max") %>%
  ggplot(aes(x = rainfall, y = temperature)) +        
  geom_point(size = 2) + 
  geom_smooth(method = "lm") +        
  labs(title = "Relationship between rainfall and max temperature") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

What about the relationship between rainfall and min temperature:

```{r}
combined_data %>%
  filter(city_name == "PERTH", year == 2014, temp_type == "min") %>%
  ggplot(aes(x = rainfall, y = temperature)) +        
  geom_point(size = 2) + 
  geom_smooth(method = "lm") +        
  labs(title = "Relationship between rainfall and min temperature") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

**Findings**

Although the relationship we see in both cases is not clear and strong, we do see a week positive association between rainfall and temperature. However, this result might be biased by the outliers. 